---
title: "Minimum Response Rate is a Crucial Part of the Polling Procedures"
subtitle: "Traffic-Calming-related polls are still consistently invalidated by the amended minimum response rate in the City of Toronto"
author: 
  - Tian Yi Zhang
thanks: "Code and data supporting this paper are available at: https://github.com/zhan7818/MRR_in_polls_City_of_Toronto."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Polls are a widely used and effective way for administrative divisions to determine the general opinion of residents and businesses on various topics. We obtain the polls conducted by the city of Toronto broken down based on the type of application and analyze it using two graphs and a table. The data shows a consistent trend for parking-related polls to meet the required engagement rate, while the opposite is seen for traffic calming applications. The findings suggest a change in percentage for the number of returned ballots required for polls regarding traffic calming."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor) # for clean_names()
library(knitr) # for kable()
```

# Introduction

What is the best way to find out what people think about a particular topic?
Ask them. This sentiment is shared
by Joshua Clinton from the Vanderbilt University [@citevanderbilt], who believes
that "public opinion polling, done right, remains the best way of obtaining
citizens' opinions." While he does raise the point that polls should not be
the only thing that public policy is based on, it is nevertheless a good way to
gauge the public's opinions and expectations. This is the crux of the concept 
of polling, and is why it is a widely employed tactic.
As an administrative body, the City of Toronto is no
stranger to utilizing polls to determine the
opinions of its residents and business owners. These polls could be for any
topic, ranging from changes to by-laws, to a simple addition of a speed hump
on a particular road, and have no doubt aided the City of Toronto in
constructing improvements to the city.

An important factor in conducting a poll is the idea
of the $\textbf{minimum response rate}$, which could be described as 
the percentage
of ballots that need to be returned in order for the poll to be considered
valid. The City of Toronto initially adopted a minimum response rate of 50% 
in 2006 [@citebackground37202], and while many polls were considered valid
under this 50% rate, it is also true that a significant portion of polls were
invalidated because of it. This would not be an issue on its own, but when
the invalidated polls often have an overwhelmingly positive response, as is
illustrated by Councilor Frances Nunziata [@citebackground35072], there is
a case to be made for the argument that perhaps the minimum response rate
of 50% has invalidated many useful polls and is therefore too high. The
minimum response rate was eventually lowered down to 25% after many reviews
[@citebackground37202], and it is the primary goal of this paper to examine
whether this change was a meaningful one.

This paper will be using the dataset for polls conducted by the City of Toronto
from 2015 to 2021, obtained from City of Toronto's Open Data Portal,
to analyze the relationship between the minimum response rate
and the number of ballots cast by the voters to see if the change made to
the minimum response rate had any impact on the validity of polls, particularly
the ones concerned with Front Yard Parking and Traffic Calming. The results
from this analysis will have clear implications for further changes towards
the minimum response rate for polls conducted by the City.

The remainder of this paper contains Section \@ref(data), which
explains the sources of the dataset, how it is structured, its strengths,
and its weaknesses. The findings from observations on the graphs and table are also
included in this section.

# Data

The dataset used in this paper is obtained from
the City of Toronto open data portal, using the `opendatatoronto` package
[@citeopendatatoronto] and the statistical programming language R [@citeR].
A new variable, `ballots_needed_to_proceed_double`, will be constructed by
doubling the number of ballots needed to proceed, `ballots_needed_to_proceed`,
to simulate the minimum response rate before the amendment from 50% to 25%
as discussed in Development and Reviews of Front Yard Parking Poll Requirements
[@citebackground37202]. This newly constructed variable is used in
Figure \@ref(fig:graph).

In addition to using R [@citeR] to construct this paper, several R packages are used
as well: The feature names in the dataset are cleaned using the `janitor` package
[@citejani]. Subsequent graphs and tables shown in this paper also utilize the
`tidyverse` [@citetidy] and `knitr` [@citeknitr] package respectively.

The dataset is collected by the Toronto City Clerk's Office,
whom are responsible for administering polls and collecting the results. The
data collection started from as early as April 1, 2015, and is updated daily
after each closing and certification of a poll.
The polls included in this dataset are used to determine whether property owners
and businesses have a favorable attitude towards a change in their neighbourhood
that could affect them. The types of polls are divided into the following five
categories:

* Boulevard Cafes
* Off-Street Parking (such as front yard parking and commercial boulevard parking)
* Permit Parking
* Traffic Calming
* Business Improvement Area

The majority of the polls conducted in this dataset are parking and traffic-calming
related, with the rest of the polls dealing with public opinion concerning business 
improvement areas and boulevard cafes.

According to information gathered from chapter 190 of City of Toronto bylaws 
[@citechp] and the City of Toronto webpage regarding polls [@citeopenpoll],
the procedure of 
polling can be summarized as follows: First, the City reviews a proposal for
an application and decides whether to approve it. If approved, the City
conducts a poll of the population in the area affected by the
application. Every owner, resident and tenant at least 18 years old in the
affected area will be mailed the materials necessary to engage in the poll,
including but not limited to: a notice regarding the poll and its deadline,
the ballot distributed, and a return envelope. Then, the results from the 
poll are collected. Note that for a poll to be considered positive, it must 
first reach a certain percentage of ballots returned so as to ensure the poll is
an accurate depiction of the general consensus in the area of interest. 
This threshold differs for each type of poll as determined by specific 
by-laws and/or city policies. Lastly, should the result of the poll be 
positive, the application can then proceed through the approval process 
(Final approval by City Council may be required based on the type of poll).

The dataset contains several features that are variations of other features
with additional details that may otherwise be useful for a different analysis.
For example, the dataset records both the number of ballots distributed,
`BALLOTS_DISTRIBUTED`, and the number of ballots returned to City Clerk's Office,
`BALLOTS_RECEIVED_BY_VOTERS`. The difference in the data recorded between these
two variables are minor or nonexistent and would perhaps be useful in another
analysis regarding the element of human error present in the execution of polling
procedures. Other minor but potentially crucial details, such as the number of
ballots received but not clearly marked as "in favour" or "opposed", recorded as
`BALLOTS_SPOILED`, are also included in the dataset. These small details included
in the dataset allow for more accurate data analysis depending on the topic
presented in a paper, and speaks volumes about how comprehensive the dataset
is.

That is not to say that the dataset is flawless, however. Certain issues 
arise with the aforementioned polling methods, too: Despite
stating that every eligible voter in the affected area will be able to
participate in the poll, it is not guaranteed that this will be the case.
Table \@ref(tab:dataextract) showcases this discrepancy:

```{r dataextract, echo=FALSE, warning=FALSE, message=FALSE}
# Read in the dataset
polls_conducted_by_city <- 
  read_csv(here::here("inputs/data/polls_conducted_by_city.csv"))

# Clean the dataset
polls_clean <- polls_conducted_by_city |>
  clean_names()

# Create a new variable that doubles the number of ballots required to proceed
polls_clean <- polls_clean |>
  mutate(ballots_needed_to_proceed_double = ballots_needed_to_proceed * 2)

# Create a table that showcases the dataset
polls_clean |> 
  select(address, potential_voters, ballots_distributed, ballots_cast) |>
  drop_na() |>
  slice(1:10) |> 
  kable(
    caption = "First ten rows of a dataset that shows poll engagement
    in Toronto from 2015 to 2021",
    col.names = c("Address","Potential Voters","Ballots Distributed","Ballots Cast"),
    digits = 1,
    booktabs = TRUE,
    linesep = ""
  )
```

Note that the number of potential voters differs from the number of ballots
distributed when they should ideally be equal. This difference could be due to
human errors or is intentional, which means that
even if all ballots distributed are properly returned, the results are not
representative of the true population. Alternatively, this discrepancy could be
due to the way the data is collected and labelled: The City of Toronto's
[@citeopendatatoronto]
official description of `POTENTIAL_VOTERS`, the feature in the dataset which
records the number of potential voters, states
that this feature represents the "number of people residing within poll
boundary range." However, as we previously stated, there are certain
requirements that must be fulfilled for one to be considered eligible for the poll
(for example, all participants must be at least 18 years of age). The vagueness
in the description and naming of the `POTENTIAL_VOTERS` variable means that
we are not sure if residents under the age of 18 in the affected area
are also included in the dataset
as a "potential voter", and regrettably cannot accurately assess the reasoning 
behind the discrepancy between
the number of potential voters and the number of ballots distributed.

Assuming that the first reasoning is true, then the dataset may be biased in
several possible ways depending on how the ballots are distributed: Recall from
chapter 190 of the City of Toronto bylaws [@citechp] that the City Clerk's
Office is responsible for compiling the polling list which includes the names
of the participants and their addresses. Is it possible that a racial bias may
be present in the dataset as the City Clerk's Office intentionally omits
affected residents of a certain minority group? Or perhaps the City Clerk's
Office is prone to contextual bias as they get to decide which residents
are considered affected by the application proposal and which ones are not?
There is no way to
analyze these possibility because relevant data such as the ethnicity of
affected residents are not included in the dataset, a flaw that we should not
ignore when examining the reliability of the dataset.

The next primary subject of interest is the relationship between the number of
ballots cast, the number of ballots needed for the poll to be considered 
valid, and the type of application. This relationship could be used as a measure
of how often a poll is invalidated by the minimum response rate, separated
depending on the type
of application, and affects future choices for setting new minimum
response rates. Figure \@ref(fig:graph) shows
the relationship between the number of ballots
returned and the number of ballots needed to proceed
$\textbf{before the reduction of minimum response rate from 50\% to 25\%}$
as discussed in
Development and Reviews of Front Yard Parking Poll Requirements
[@citebackground37202].

```{r graph, fig.cap="Ballots distributed versus ballots cast for each type of application, 50% return requirement", echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
# make a graph. row 182 with id=800776 is filtered out because it is very far
# away from the rest of the dataset and squishes the rest of the dataset,
# resulting in a poor graph viewing experience
# A line was added to help readers see which polls are not considered valid
# due to having less ballots cast than the required number
polls_clean[-c(182),] |>
  ggplot(aes(x = ballots_cast,
             y = ballots_needed_to_proceed_double,
             color = application_for)) +
  geom_point(alpha = 0.8) +
  geom_abline(intercept=0, slope=1, size=1, alpha=0.3) +
  labs(y = "Number of Ballots Needed to Proceed",
       x = "Number of Ballots Cast")
```

We can see from Figure \@ref(fig:graph) that the majority of the polls from
this dataset are for traffic calming and parking applications.
Note that the graph shows the
majority of parking-related polls are close to each other on the graph and
generally have less than 100 ballots cast.
This makes sense because applications concerning parking
only affects the residents of the property being approved for parking.
In Contrast, polls regarding traffic calming vary much more in terms of number
of ballots cast, which makes sense because improvements to certain
road sections would affect more people than the rest.

The black line represents the line where the ballots cast is equal to the
required number of ballots needed to proceed; the points above the black line
represent polls that did not meet the required number of returned ballots needed
to be considered valid. Note that in Figure \@ref(fig:graph),
the number of required returned ballots for a
poll to be considered valid is 50% of the number of ballots distributed.
We can see that Councillor Frances Nunziata's [@citebackground35072] 
concerns are supported by this dataset; There was not just a few, but a significant
number of Front Yard Parking polls considered invalid polls under the 50%
minimum response rate policy.

The previously mentioned percentage of 50 was subject to many discussions
and revisions, as noted in Development and Reviews of Front Yard Parking
Poll Requirements [@citebackground37202], to the point where it is now reduced from
50% to 25%, the latter shown in Figure \@ref(fig:graph2).


```{r graph2, fig.cap="Ballots distributed versus ballots cast for each type of application, 25% return requirement", echo = FALSE, fig.width=8, fig.height=4, warning=FALSE, message=FALSE}
# make a graph. row 182 with id=800776 is filtered out because it is very far
# away from the rest of the dataset and squishes the rest of the dataset,
# resulting in a poor graph viewing experience
# A line was added to help readers see which polls are not considered valid
# due to having less ballots cast than the required number
polls_clean[-c(182),] |>
  ggplot(aes(x = ballots_cast,
             y = ballots_needed_to_proceed,
             color = application_for)) +
  geom_point(alpha = 0.8) +
  geom_abline(intercept=0, slope=1, size=1, alpha=0.3) +
  labs(y = "Number of Ballots Needed to Proceed",
       x = "Number of Ballots Cast")
```

Observations from Figure \@ref(fig:graph2) tells us that most of the 
applications for parking-related applications
meet the amended minimum response rate of 25%. This,
in conjunction with our observation from Figure \@ref(fig:graph), shows
that the issues brought up and reviewed by City Council [@citebackground37202],
particularly Councillor Frances Nunziata [@citebackground35072], were valid
concerns regarding the polling procedures, and that the reduction of
minimum response rate from 50% to 25% was a step in the right direction.

Another important piece of information presented by Figure \@ref(fig:graph2)
is that despite the reduction in the required number of ballots returned,
a significant portion of polls regarding traffic calming still lies above
the black line, failings to meet the 25% minimum response rate. It would be
bizarre to attribute this observed consistent invalidation of polls to a simple
lack of engagement
from the affected population. Therefore, this observation implies
that there may be a need for further reduction in minimum response rate
for the poll validity requirement.

\newpage


# References


